A novel method for detecting coalescing binaries in near realtime with Advanced LIGO and beyond \editorial{Submit to PRD or CQG?}

### 2011-02-05 introduction.tex
Small fixes and comments
git-svn-id: https://svn.ligo.caltech.edu/svn/6th-Floor-Papers/lloid_paper@601 ccdd05d5-56c1-461a-bf98-85f6876523b4

As the gravitational wave detection horizon pushes outward, the ability to detect signals in near realtime will become valuable.  Having an electromagnetic or neutrino counterpart to a 
gravitational-wave detection would not only increase the confidence in 
the detection but will also greatly improve the astrophysical information
available from the event.  Most models that predict simultaneous
gravitational-wave and electromagnetic observations also predict that the
peak amplitude of electromagetic radiation will occur soon after 
gravitational-wave emmission~\cite{sylvestre2003}.  Thus in order to maximize
the chance of a successful electromagnetic followup the latency of 
gravitational-wave signal analysis must be made minimal.

\editorial{This is old news, and doesn't seem relevant to me: EM-directed gravitational wave searches don't benefit from low-latency detection pipelines.} Work has already commenced to trigger gravitational-wave searches from 
electromagnetic observations~\cite{triggeredsearches2008}. \textsc{ligo} observations
ruled out the possibility of GRB70201 originating from the merger of a
\textsc{nsns} or \textsc{nsbh} system in the Andromeda galaxy~\cite{GRB070201}.

@@

This question has been investigated and it seems  at least
possible that gravitational-wave detections could provide a region of the
sky to prompt electromagnetic observation followups~\cite{sylvestre2003}.
\editorial{Cite an in-preparation paper of Larry's?} Much work is currently underway in providing the best possible source 
localization from gravitational-wave detector networks~\cite{markowitz2008, raymond2008, cavalier2006}.  Colloborations are currently forming to provide infrastructure for the gravitational-wave
Astronomers to provide targets of opportunity for electromagnetic 
astronomers~\cite{kanner2008}.  Most models that predict coincident
gravitational-wave and electromagnetic observations also predict that the
peak electromagetic fluence will occur soon after 
gravitational-wave emission~\cite{sylvestre2003}.  Thus in order to maximize
the chance of a successful electromagnetic followup the latency of 
gravitational-wave signal analysis must be made minimal.

>>>>

As the gravitational wave detection horizon pushes outward, the ability to detect signals in near realtime will become valuable.  Having an electromagnetic or neutrino counterpart to a 
gravitational-wave detection would not only increase the confidence in 
the detection but will also greatly improve the astrophysical information
available from the event.  Most models that predict simultaneous
gravitational-wave and electromagnetic observations also predict that the
peak amplitude of electromagetic radiation will occur soon after 
gravitational-wave emission~\cite{sylvestre2003}.  Thus in order to maximize
the chance of a successful electromagnetic followup the latency of 
gravitational-wave signal analysis must be made minimal.

\editorial{This is old news, and doesn't seem relevant to me: EM-directed gravitational wave searches don't benefit from low-latency detection pipelines.} Work has already commenced to trigger gravitational-wave searches from 
electromagnetic observations~\cite{triggeredsearches2008}. \textsc{ligo} observations
ruled out the possibility of GRB070201 originating from the merger of a
\textsc{nsns} or \textsc{nsbh} system in the Andromeda galaxy~\cite{GRB070201}.

@@

This question has been investigated and it seems  at least
possible that gravitational-wave detections could provide a region of the
sky to prompt electromagnetic observation followups~\cite{sylvestre2003}.
\editorial{Cite an in-preparation paper of Larry's?} Much work is currently underway in providing the best possible source 
localization from gravitational-wave detector networks~\cite{markowitz2008, raymond2008, cavalier2006}.  Collaborations are currently forming to provide infrastructure for the gravitational-wave
Astronomers to provide targets of opportunity for electromagnetic 
astronomers~\cite{kanner2008}.  Most models that predict coincident
gravitational-wave and electromagnetic observations also predict that the
peak electromagetic fluence will occur soon after 
gravitational-wave emission~\cite{sylvestre2003}.  Thus in order to maximize
the chance of a successful electromagnetic followup the latency of 
gravitational-wave signal analysis must be made minimal.

### 2011-02-05 method.tex
Small fixes and comments
git-svn-id: https://svn.ligo.caltech.edu/svn/6th-Floor-Papers/lloid_paper@601 ccdd05d5-56c1-461a-bf98-85f6876523b4

Much more commonly, the matched filter bank is implemented using \textsc{fft} convolutions, costing only $\mathcal O(M \lg N)$ per sample but having a latency of at least $M$ samples.  For example, for a bank of $1\,\mathrm{ks}$ templates sampled at $4096\,\mathrm{Hz}$, the \textsc{fft} implementation requires about about $2.2 \times 10^4$ times fewer floating point operations per sample than the \textsc{fir} implementation.  However, the \textsc{fft} implementation has a latency of at least $1\,\mathrm{ks}$.

@@

\editorial{The time slice boundary notation is a little awkward and different identifiers are used in a few places in the source code.  It would be nice to clean that up.}  It is possible to strike a balance between the low latency of the \textsc{fir} filter method and the low overhead of the \textsc{fft} method by breaking the templates into disjoint time intervals, or {\em time slices}.  The outputs of the filters for all of the time slices are delayed by the appropriate numbers of samples and then summed to produce the \textsc{snr}s for original templates in the template bank.  Each time slice may be implemented using either \textsc{fir} filters or \textsc{fft} convolution.  If the $i$th time slice spans the times $[-t_\mathrm{end}^i, -t_\mathrm{start}^i)$ relative to the time of coalescence, then the latency is $\max \, (2 (t_\mathrm{end}^i - t_\mathrm{start}^i) - t_\mathrm{start}^i)$, where the maximum is taken over all time slices that are implemented using \textsc{fft} convolution.

>>>>

Much more commonly, the matched filter bank is implemented using \textsc{fft} convolutions, costing only $\mathcal O(M \lg N)$ \editorial{$\lg$ is $\log_2$. Explain nomenclature? Switch to $\log_2$?} per sample but having a latency of at least $M$ samples.  For example, for a bank of $1\,\mathrm{ks}$ templates sampled at $4096\,\mathrm{Hz}$, the \textsc{fft} implementation requires about about $2.2 \times 10^4$ times fewer floating point operations per sample than the \textsc{fir} implementation.  However, the \textsc{fft} implementation has a latency of at least $1\,\mathrm{ks}$.

@@

\editorial{The time slice boundary notation is a little awkward and different identifiers are used in a few places in the source code.  It would be nice to clean that up.}  It is possible to strike a balance between the low latency of the \textsc{fir} filter method and the low overhead of the \textsc{fft} method by breaking the templates into disjoint time intervals, or {\em time slices}.  The outputs of the filters for all of the time slices are delayed by the appropriate numbers of samples and then summed to produce the \textsc{snr}s for original templates in the template bank.  Each time slice may be implemented using either \textsc{fir} filters or \textsc{fft} convolution.  If the $i$th time slice spans the times $[-t_\mathrm{end}^i, -t_\mathrm{start}^i)$ relative to the time of coalescence, then the latency is $\max \, (2 (t_\mathrm{end}^i - t_\mathrm{start}^i) - t_\mathrm{start}^i)$, where the maximum is taken over all time slices that are implemented using \textsc{fft} convolution.
\editorial{It may be worth pointing out that time-slicing buys you latency even
without exploiting time-frequency structure. The downsampling really just buys
you throughput. It may even be worth renaming the section.}
